---
output: github_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README",
  out.width = "100%"
)
#
options(crayon.enabled = TRUE, crayon.colors = 256)
crayon::num_colors(forget = TRUE)
asciicast::init_knitr_engine(
  startup = quote({
    library(waldo)
    conflicted::conflict_prefer("compare", "waldo")
  }),
  echo = TRUE,
  echo_input = FALSE
)
set.seed(1014)
```

# waldo

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/r-lib/waldo/branch/master/graph/badge.svg)](https://codecov.io/gh/r-lib/waldo?branch=master)
[![R build status](https://github.com/r-lib/waldo/workflows/R-CMD-check/badge.svg)](https://github.com/r-lib/waldo/actions)
<!-- badges: end -->

The goal of waldo is to find and concisely describe the difference between a pair of R objects, with the primary goal of making it easier to figure out what's gone wrong in your unit tests. 

`waldo::compare()` is inspired by `all.equal()`, but takes additional care to generate actionable insights by:

* Ordering the differences from most important to least important.
* Displaying the values of atomic vectors that are actually different.
* Carefully using colour to emphasise changes (while still being readable
  when colour isn't available).
* Using R code (not a text description) to show where differences arise.
* Where possible, comparing elements by name, rather than by position.
* Erring on the side of producing too much output, rather than too little.

## Installation

You can install the released version of waldo from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("waldo")
```

## Comparisons

```{r setup}
library(waldo)
```

When comparing atomic vectors, `compare()` produces diffs (thanks to [diffobj](https://github.com/brodieG/diffobj)) that highlight additions, deletions, and changes, along with a little context:

*   Addition
    ```{asciicast}
    compare(c("a", "b", "c"), c("a", "b"))
    ```

*   Deletion
    ```{asciicast}
    compare(c("a", "b"), c("a", "b", "c"))
    ```

*   Change
    ```{asciicast}
    compare(c("a", "b", "c"), c("a", "B", "c"))
    ```

*   Long vectors with short differences only show local context around
    changes, not everything that's the same.

    ```{asciicast}
    compare(c("X", letters), c(letters, "X"))
    ```

Depending on the relative size of the differences and the width of your console you'll get one of three displays:

*   The default display is to show the vectors one atop the other:

    ```{asciicast}
    compare(letters[1:5], letters[1:6])
    ```

*   If there's not enough room for that, the two vectors are shown 
    side-by-side:

    ```{asciicast}
    options(width = 20)
    compare(letters[1:5], letters[1:6])
    ```

*   And if there's still not enough room for side-by-side, the each element
    is given its own line:

    ```{asciicast}
    options(width = 10)
    compare(letters[1:5], letters[1:6])
    ```

When comparing more complex objects, waldo creates an executable code path telling you where the differences lie:

```{asciicast, include = FALSE}
options(width = 80)
```

*   Unnamed lists are compared by position: 

    ```{asciicast}
    compare(list(factor("x")), list(1L))
    ```

*   Named lists, including data frames, are compared by name. For example,
    note that the following comparison reports a difference in the class and
    names, but not the values of the columns.

    ```{asciicast}
    df1 <- data.frame(x = 1:3, y = 3:1)
    df2 <- tibble::tibble(rev(df1))
    compare(df1, df2)
    ```

*   Recursion can be arbitrarily deep:

    ```{asciicast}
    x <- list(a = list(b = list(c = list(structure(1, e = 1)))))
    y <- list(a = list(b = list(c = list(structure(1, e = "a")))))
    compare(x, y)
    ```
